Gaussian variance estimation in simulated data sets

### 2018-11-09 analysis/gaussvarest.Rmd
wflow_publish("gaussvarest.Rmd")

```{r knitr-opts, include=FALSE}
knitr::opts_chunk$set(comment = "#",collapse = TRUE,fig.align = "center")
```

@@

  mse.mu.uneven.smash[j] <- mean((mu.est.inter-fTrue(xgOrig))^2)
  mse.sd.uneven.smash[j] <-
    mean((sqrt(var.est.inter)-exp((loggTrue(xgOrig))/2))^2)
}
cat("\n")
```

@@

In this scenario, we simulated evenly spaced data points.

```{r, eval=FALSE}
mse.mu.even.mfvb <- 0
mse.mu.even.s8   <- 0

mse.sd.even.mfvb <- 0
mse.sd.even.s8   <- 0
lmse.sd.even.mfvb <- 0
lmse.sd.even.s8   <- 0

# Repeat for each data set simulated in the second simulation scenario.
cat(sprintf("Running %d simulations: ",nsim2))
for (j in 1:nsim2) {
  cat(sprintf("%d ",j))

  # Create the simulated data set
  # -----------------------------
  n      <- 2^10
  xOrig  <- (1:n)/n
  set.seed(30*j)
  yOrig  <- fTrue(xOrig) + sqrt(exp(loggTrue(xOrig)))*rnorm(n)
  aOrig  <- min(xOrig)
  bOrig  <- max(xOrig)
  mean.x <- mean(xOrig)
  sd.x   <- sd(xOrig)
  mean.y <- mean(yOrig)
  sd.y   <- sd(yOrig)

@@

  # Run Mean Field Variational Bayes (MFVB)
  # ---------------------------------------
  X     <- cbind(rep(1,n),x)
  Cumat <- cbind(X,Zu)
  Cvmat <- cbind(X,Zv)
  ncX   <- ncol(X)
  ncZu  <- ncol(Zu)
  ncZv  <- ncol(Zv)
  ncCu  <- ncol(Cumat)
  ncCv  <- ncol(Cvmat)

@@

  # Run SMASH
  # ---------
  mu.est  <- smash.gaus(yOrig,filter.number=1,family="DaubExPhase")
  var.est <- smash.gaus(yOrig,v.est=TRUE)

  mse.mu.even.mfvb[j] <- mean((fhatMFVBgOrig-fTrue(xgOrig))^2)
  mse.mu.even.haar[j] <- mean((mu.est-fTrue(xgOrig))^2)

  mse.sd.even.mfvb[j]  <- mean((sqrtghatMFVBgOrig-exp((loggTrue(xgOrig))/2))^2)
  mse.sd.even.haar[j]  <- mean((sqrt(var.est)-exp((loggTrue(xgOrig))/2))^2)
  lmse.sd.even.mfvb[j] <- mean((log(sqrtghatMFVBgOrig)-(loggTrue(xgOrig))/2)^2)
  lmse.sd.even.haar[j] <- mean((log(sqrt(var.est))-(loggTrue(xgOrig))/2)^2)

  mu.est       <- smash.gaus(yOrig,filter.number=8,family="DaubLeAsymm")
  var.est      <- smash.gaus(yOrig,v.est=TRUE,v.basis=TRUE,filter.number=8,
                             family="DaubLeAsymm")
  var.est.s8.j <- smash.gaus(yOrig,v.est=TRUE,v.basis=TRUE,jash=TRUE,
                             filter.number=8,family="DaubLeAsymm")
  var.est.j    <- smash.gaus(yOrig,v.est=TRUE,jash=TRUE)

  mse.mu.even.s8[j]   <-mean((mu.est-fTrue(xgOrig))^2)
  mse.sd.even.s8[j]   <-mean((sqrt(var.est)-exp((loggTrue(xgOrig))/2))^2)
  mse.sd.even.j[j]    <-mean((sqrt(var.est.j)-exp((loggTrue(xgOrig))/2))^2)
  mse.sd.even.s8.j[j] <-mean((sqrt(var.est.s8.j)-exp((loggTrue(xgOrig))/2))^2)
  lmse.sd.even.s8[j]  <-mean((log(sqrt(var.est))-(loggTrue(xgOrig))/2)^2)
  lmse.sd.even.j[j]   <-mean((log(sqrt(var.est.j))-(loggTrue(xgOrig))/2)^2)
  lmse.sd.even.s8.j[j]<-mean((log(sqrt(var.est.s8.j))-(loggTrue(xgOrig))/2)^2)
}
cat("\n")

# SUMMARIZE RESULTS
# -----------------
# Compare this table against Table 1 in the paper.
cat("Summarizing results of simulations.\n")
cat(sprintf("MSE averaged across %d simulations in Scenario 1,\n",nsim1))
cat(sprintf("and averaged across %d simulations in Scenario 2:\n",nsim2))
mse.table <- rbind(c(mean(mse.mu.uneven.mfvb),mean(mse.sd.uneven.mfvb),
                     mean(mse.mu.even.mfvb),mean(mse.sd.even.mfvb)),
                   c(mean(mse.mu.uneven.s8),mean(mse.sd.uneven.s8),
                     mean(mse.mu.even.s8),mean(mse.sd.even.s8)))
rownames(mse.table) <- c("MFVB","SMASH")
colnames(mse.table) <- c("mean","sd","mean","sd")
cat("         Scenario 1      Scenario 2   \n")
cat("      --------------- ----------------\n")
print(mse.table)
```

>>>>

```{r knitr-opts, include=FALSE}
knitr::opts_chunk$set(comment = "#",collapse = TRUE,results = "hold",
                      fig.align = "center")
```

@@

  mse.mu.uneven.smash[j] <- mean((mu.est.inter-fTrue(xgOrig))^2)
  mse.sd.uneven.smash[j] <-
    mean((sqrt(var.est.inter)-exp((loggTrue(xgOrig))/2))^2)
}
```

@@

In this scenario, we simulate data sets with 1,024 evenly spaced data
points. We assess accuracy separately for the mean and standard
deviation as the mean of the MSEs evaluated at each of the locations.

```{r simulation-2-initialize}
mse.mu.even.mfvb  <- 0
mse.mu.even.smash <- 0
mse.sd.even.mfvb  <- 0
mse.sd.even.smash <- 0
```

```{r simulation-2-main-loop}
cat(sprintf("Running %d simulations: ",nsim2))
for (j in 1:nsim2) {
  cat(sprintf("%d ",j))

  # SIMULATE DATA
  n      <- 2^10
  xOrig  <- (1:n)/n
  set.seed(30*j)
  yOrig  <- fTrue(xOrig) + sqrt(exp(loggTrue(xOrig)))*rnorm(n)
  aOrig  <- min(xOrig)
  bOrig  <- max(xOrig)
  mean.x <- mean(xOrig)
  sd.x   <- sd(xOrig)
  mean.y <- mean(yOrig)
  sd.y   <- sd(yOrig)

@@

@@

  # RUN SMASH
  mu.est  <- smash.gaus(yOrig,filter.number=1,family="DaubExPhase")
  var.est <- smash.gaus(yOrig,v.est=TRUE)

  mse.mu.even.mfvb[j] <- mean((fhatMFVBgOrig-fTrue(xgOrig))^2)
  mse.sd.even.mfvb[j] <- mean((sqrtghatMFVBgOrig-exp((loggTrue(xgOrig))/2))^2)

  mu.est  <- smash.gaus(yOrig,filter.number=8,family="DaubLeAsymm")
  var.est <- smash.gaus(yOrig,v.est=TRUE,v.basis=TRUE,filter.number=8,
                        family = "DaubLeAsymm")

  mse.mu.even.smash[j]  <- mean((mu.est - fTrue(xgOrig))^2)
  mse.sd.even.smash[j]  <- mean((sqrt(var.est)-exp((loggTrue(xgOrig))/2))^2)
}
```

Summarize results of simulations
--------------------------------

Compare this table against Table 1 in the paper.

```{r summarize-mse}
cat(sprintf("MSE averaged across %d simulations in Scenario 1,\n",nsim1))
cat(sprintf("and averaged across %d simulations in Scenario 2:\n",nsim2))
mse.table <- rbind(c(mean(mse.mu.uneven.mfvb),mean(mse.sd.uneven.mfvb),
                     mean(mse.mu.even.mfvb),mean(mse.sd.even.mfvb)),
                   c(mean(mse.mu.uneven.smash),mean(mse.sd.uneven.smash),
                     mean(mse.mu.even.smash),mean(mse.sd.even.smash)))
rownames(mse.table) <- c("MFVB","SMASH")
colnames(mse.table) <- c("mean","sd","mean","sd")
cat("         Scenario 1      Scenario 2   \n")
cat("      --------------- ----------------\n")
print(mse.table)
```
